---
title: "The Landscape of Interactive Visualization"
subtitle: ""  
author: "Dave Armstrong"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["xaringan-themer.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(dev = 'svg')
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE, tidy=TRUE,  tidy.opts = list(only.comment=TRUE,  width.cutoff=80))
knitr::opts_chunk$set(fig.retina=2)
knitr::opts_hooks$set(fig.callout = function(options) {
  if (options$fig.callout) {
    options$echo <- FALSE
    options$out.height <- "99%"
    #   options$fig.width <- 16
    #   options$fig.height <- 8
  }
  options
})
xaringanExtra::use_scribble(pen_color="#3252a8", pen_size=5, 
    palette=c("#e41a1c", "#4daf4a", "#ff7f00", "#4F2683", "#3252a8") 
)
xaringanExtra::use_tile_view()
xaringanExtra::use_clipboard()


library(car)
library(tibble)
library(ggplot2)
library(formatR)
library(knitr)
library(pander)
library(xtable)
library(dplyr)
library(tidyr)
library(xaringanthemer)
style_mono_accent(base_color = "#4F2683", code_font_size=".65rem", text_font_size = "1.25rem")
library(glue)
library(plotly)
library(psre)
library(rio)
library(daviz)
library(chorddiag)
library(sf)
library(countrycode)
library(crosstalk)
library(leaflet)
library(DT)
data(wvs)
ifWidget <- function(plot, h=500, w=500, file=1, return=TRUE, selfcontained=TRUE){
  if(is.numeric(file)){
    f <- paste0("p.", file, ".html")  
  }else{
    f <- file
  }
  htmlwidgets::saveWidget(plot, f, selfcontained=selfcontained)
  if(return){
  htmltools::tags$iframe(
    src=f,
    height=as.character(h),
    width=as.character(w),
    scrolling="no", 
    seamless="seamless", 
    frameBorder="0")
  }
}

```

```{css, echo=FALSE}
/* custom.css */
.left-code {
  color:
    
   
    
    #  777;
  width: 35%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 63%;
  float: right;
  padding-left: 1%;
}
.right-plot-shift {
  width: 63%;
  float: right;
  padding-left: 1%;
  position:relative; 
  top: -100px;
}
.right-plot-shift2 {
  width: 63%;
  float: right;
  padding-left: 1%;
  position:relative; 
  top: -50px;
}
.right-plot-shift3 {
  width: 63%;
  float: right;
  padding-left: 1%;
  position:relative; 
  top: -25px;
}
.shift { 
  position:relative; 
  top: -100px;
  }

.shift150 { 
  position:relative; 
  top: -150px;
  }


.plot-callout {
  height: 225px;
  width: 450px;
  bottom: 5%;
  right: 5%;
  position: absolute;
  padding: 0px;
  z-index: 100;
}
.plot-callout img {
  width: 100%;
  border: 4px solid
  #  23373B;
}

.pull-right-shift {
  float: right;
  width: 47%;
  position: relative; 
  top: -100px;
}

.pull-right-shift2 {
  float: right;
  width: 47%;
  position:relative; 
  top: -50px;
}


.pull-right ~ * {
  clear: both;
}


.mycol {
  float: left;
  width: 30%;
  padding: 5px;
}

/* Clear floats after image containers */
.myrow::after {
  content: "";
  clear: both;
  display: table;
}

```

```{r eval=FALSE}
remotes::install_github("davidaarmstrong/daviz", ref= "f2addeeac7496a5a85ddbaaf7f59f515fba9a973")
remotes::install_github("mattflor/chorddiag")
```


## Interactive Displays as HTML (Client-side Computing). 
1. Introduction
1. What are the possibilities? 
  - Points/Lines/Coefficients, Surfaces, Maps, Chords
2. Software for Web Development. 
3. Software for Making Interactive Plots. 
4. Dashboards and Apps. 
5. Hosting html-only interactive elements. 

---

## Introduction

Interactive graphics allow users (to some degree) to control and interact with a visual display. 

- Client-side computing
  - Renders directly in the browser (using Javascript). 
  - Can be hosted on any web server. 
- Server-side computing   
  - Requires other non-client-side software (e.g., R, Python, etc...) to do computations on the fly in response to user inputs.
  - Must be hosted in a special environment (e.g., shiny server). 
---


## Scatterplots

<center>

```{r, "plotly_points"}
g <- wvs %>% 
  filter(wave == 5) %>% 
  mutate(txt = glue("Country: {country}\nGINI: {round(gini_disp,2)}",
                    "\nEmancipative Values: {round(resemaval, 2)}")) %>% 
  ggplot(aes(x=gini_disp, y=resemaval, text = txt)) + 
    geom_point() + 
  theme_classic() + 
  labs(x="Gini (disposable income)", y="Emancipative Values")
ggplotly(g, tooltip = "text")
```

</center>

---

## Points (Different Series)

<center>
```{r "plotly_points_colour", out.width="80%"}
g1 <- wvs %>% 
  filter(wave == 5) %>% 
  mutate(txt = glue("Country: {country}\nGINI: {round(gini_disp,2)}",
                    "\nSecular Values: {round(sacsecval, 2)}"), 
         Democrat = factor(democrat, 
                            levels= 1:2, 
                            labels =c("Non-Democracy", "Democracy"))) %>% 
  ggplot(aes(x=gini_disp, y=sacsecval, text = txt, colour=Democrat)) + 
    geom_point() + 
    theme_classic() + 
    labs(x = "Gini (disposable income)", y="Emancipative Values")
ggplotly(g1, tooltip = "text")
```
</center>


---


## Coefficient Plots

<center>
```{r, "coefd3", results="asis"}
w2 <- wvs %>% 
  filter(wave == 5) %>% 
  mutate(txt = glue("Country: {country}\nGINI: {round(gini_disp,2)}",
                    "\nSecular Values: {round(sacsecval, 2)}"), 
         Democrat = factor(democrat, 
                            levels= 1:2, 
                            labels =c("Non-Democracy", "Democracy")), 
         civ_c = case_when(civ == 1 ~ "African", 
                           civ == 4 ~ "Islamic", 
                           civ == 6 ~ "Latin American", 
                           civ == 7 ~ "Orthodox", 
                           civ == 9 ~ "Western", 
                           TRUE ~ "Other"), 
         Civilization = as.factor(civ_c)) 
mod2 <- lm(resemaval ~ Civilization, data=w2)
sigd3(mod2, "Civilization", "civ.html")
```

</center>

---


## Lines
<center>
```{r "plotly_lines", out.width="100%"}
g3 <- wvs %>% 
  group_by(country) %>% 
  mutate(n = n()) %>% 
  filter(n == 4) %>% 
  mutate(txt = glue("Country: {country}", 
                     "\nWave: {wave}", 
                     "\nEmancipative Values: {round(resemaval, 2)}"), 
         wave=as.numeric(wave)) %>% 
  ggplot(aes(x=wave, y=resemaval, colour=country, text=txt, group=country)) + 
  geom_line() + 
  theme_classic() + 
  labs(x="Wave", y="Emancipative Values")
ggplotly(g3, tooltip="text")

```
</center>

---


## Chords
<center>
```{r, "chords", out.width="75%", results="asis"}
load("data/ash.rda")
ash <- ash %>% 
  mutate(across(starts_with("ashcroft"), rio::factorize),
         across(starts_with("ashcroft"), as.character), 
         across(starts_with("ashcroft"), 
                ~case_when(.x %in% c("Change UK- The Independent Group", 
                                    "An independent candidate", 
                                    "British National Party (BNP)") ~ "Other",
                           TRUE ~ .x))) 
tab1 <- table(ash$ashcroftW4, ash$ashcroftW5)
tab2 <- table(ash$ashcroftW5, ash$ashcroftW12)
tabs <- tab1 + tab2
colnames(tabs) <- rownames(tabs) <- c("Con", "DK", "Grn", "Abs", "Lab", 
                                      "LD", "Other", "PC", "SNP", "UKIP")
tabs2 <- prop.table(tabs, 1)
chorddiag(tabs, 
           groupnamePadding = 30, 
           groupnameFontsize = 12, 
           groupNames = colnames(tabs),
           palette = "Set3")
```
</center>

---

## Maps

<center>
```{r "plotly_map", out.width="100%"}
load("data/fiw.rda")
data(wrld_simpl, package="maptools")
wrld_simpl <- as(wrld_simpl, "sf")
wrld_simpl <- wrld_simpl  %>% 
  mutate(cown = countrycode(ISO3, "iso3c", "cown"))
wrld_simpl <- left_join(wrld_simpl, fiw, by="cown") %>% 
  filter(!is.na(total)) %>% 
  mutate(txt = glue("Country: {country}", 
                    "\nFIW: {total}", 
                    "\nPop: {sprintf('%.1fM',POP2005/1000000)}"))
g4 <- ggplot(wrld_simpl, aes(fill=total, text=txt)) + 
  geom_sf(size=.05) + 
  scale_fill_viridis_c() + 
  labs(fill="Freedom") + 
  theme_void() + 
  theme(axis.line.x = element_blank(), 
        axis.line.y = element_blank()) 
ifWidget(ggplotly(g4), file="pmap2.html", h = 500, w=1500) 
```
</center>

---

## Other map options

- [Leaflet](https://leafletjs.com)
- [Mapbox](https://www.mapbox.com)

---

## Surfaces

```{r "plotly_surf1", cache=TRUE}
wvs3 <- wvs %>% 
  mutate(dem1 = as.numeric(democrat == 1), 
         dem2 = as.numeric(democrat == 2))
library(mgcv)
m3 <- gam(resemaval ~ 
           te(gini_disp, by=dem1) + 
           te(gdp_cap, by = dem1) + 
           ti(gini_disp, gdp_cap, by=dem1) + 
           te(gini_disp, by=dem2) + 
           te(gdp_cap, by = dem2) + 
           ti(gini_disp, gdp_cap, by=dem2),
          data=wvs3)

s_gini <- seq(min(wvs$gini_disp, na.rm=TRUE), 
         max(wvs$gini_disp, na.rm=TRUE), 
         length=25)
s_gdp_cap <- seq(min(wvs$gdp_cap, na.rm=TRUE), 
         max(wvs$gdp_cap, na.rm=TRUE), 
         length=25)

predfun <- Vectorize(function(gini, gdp, democrat = 1){
  tmp <- data.frame(gini_disp = gini, 
                    gdp_cap = gdp, 
                    dem1 = as.numeric(democrat == 1), 
                    dem2 = as.numeric(democrat == 2))
  predict(m3, newdata=tmp)
})

o1 <- outer(s_gini, s_gdp_cap, predfun, democrat = 1)
o2 <- outer(s_gini, s_gdp_cap, predfun, democrat = 2)
```
<center>
```{r "plotly_surf2"}
col1 <- rep(0, length(o1))
dim(col1) <- dim(o1)
col2 <- rep(1, length(o2))
dim(col2) <- dim(o2)

p <- plot_ly(colors = c("red", "blue")) %>% 
  add_surface(x = ~s_gini, y= ~ s_gdp_cap, z = ~o1, 
             surfacecolor=col1,
             opacity=.7,
             showscale = FALSE,
             showlegend=FALSE,
             name = "Non-Democracies",
             cauto=F,
             cmax=1,
             cmin=0) %>% 
    add_surface(x = ~s_gini, y= ~ s_gdp_cap, z = ~o2, 
             surfacecolor=col2,
             opacity=.7, 
             showscale = FALSE, 
             showlegend=FALSE,
             name = "Democracies", 
             cauto=F,
             cmax=1,
             cmin=0) %>% 
  layout(scene = list(
    xaxis = list(title="Gini"), 
    yaxis = list(title = "GDP/capita"), 
    zaxis = list(title="Emancipative Values")))
ifWidget(p, h=500, w=1000, file="surf.html")
```
</center>

---

## Software for Web Development

- Javascript (JS)
  - I took [this Udemy course](https://www.udemy.com/course/the-complete-javascript-course/). 
- Cascading Style Sheets (CSS)
  - I took [this Udemy course](https://www.udemy.com/course/design-and-develop-a-killer-website-with-html5-and-css3/). 
- HTML 
  - You should learn some of the basics of HTML.  [Here is a Udemy course](https://www.udemy.com/course/intro-to-html-web-development/) that should help. 
  

---

## Software for Interactive Graphics 

- Plotly
  - Here is the [main Plotly project page](https://plotly.com)
  - Here is the [Gallery](https://plotly.com/python/)
- Highcharts
  - Here is the [main Highcharts project page](https://www.highcharts.com/blog/tutorials/python-angular-and-highcharts/)
  - Here is a link to the [Gallery](https://www.highcharts.com/demo)
- Bokeh
  - Here is the [main Bokeh project page](https://docs.bokeh.org)
  - Here is the [Gallery](https://docs.bokeh.org/en/latest/docs/gallery.html#standalone-examples)
- D3.js
  - For the most flexibility, [D3.js](https://d3js.org) has a lot to offer, but the learning curve is much steeper. 
  - Here is the [Gallery](https://observablehq.com/@d3/gallery) 
  
---

## High-level Interaction

- Both R and Python have interfaces to the libraries mentioned above. 
  - R: [plotly](https://plotly.com/r/), [highcharter](https://jkunst.com/highcharter/), [rbokeh](https://hafen.github.io/rbokeh/), [D3.js](https://rstudio.github.io/r2d3/)
  - Python: [plotly.py](https://plotly.com/python/), [python-highcharts](https://github.com/kyper-data/python-highcharts), [Bokeh](https://bokeh.org), [D3.js examples](https://towardsdatascience.com/creating-beautiful-stand-alone-interactive-d3-charts-with-python-804117cb95a7)
- Stata has some nascent (user-written) facility with D3 graphs. 
  - [stata2d3](https://github.com/robertgrant/stata2d3)
  - [Presentation](https://www.stata.com/meeting/uk14/abstracts/materials/uk14_grant.pdf)
  - Limited [Highcharts functionality](https://github.com/bsouthga/stataHC). 
  - Could get other options through Python and R integration. 

---

## Dashboards and Apps

Both Python and R have native dashboarding capabilities (Dash and Shiny, respectively). 
  - Dash also works with R and Julia. 

Some Dashboards can be made with only client-side computing. 
  - Filtering, selecting, highlighting 
  - Requires relatively small data.
  - [Example](http://quantoid.surge.sh/apps/covid_dashboard_param.html)
  
Any dashboard or app that requires data manipulation, models, etc... require server-side computing. 
  - [Example](https://quantoid.shinyapps.io/CES_2020/)

---

## Hosting

Client-side Dashboards
  - Any web-hosting platform. 
  - [Surge](https://surge.sh) which is free, pretty easy to use and gives you a access to a free domain (e.g, `quantoid.surge.sh`). 
  
Server-side Apps and Dashboards
  - [Shinyapps.io](https://shinyapps.io) has several tiered plans (including free) depending on resource needs.
  - If money is no object, [RStudio connect](https://www.rstudio.com/products/connect/) may be the right option. 
  - [Here](https://hosting.analythium.io/how-to-pick-the-right-hosting-option-for-your-shiny-app/) is a discussion of many of the common options. 
